MGRFE: multilayer recursive feature elimination based on an embedded genetic algorithm for cancer classification

### 2018-12-01 MGRFE Response 2018-12-01/Latex Manuscript/response.txt

Dear Editor,

Enclosed please find our substantially revised manuscript “MGRFE: multilayer recursive feature elimination based on embedded genetic algorithm for cancer classification”. In this revised manuscript, we have carefully addressed all the concerns by the two reviewers. We greatly appreciate the Referee’s comments on our manuscript. The following is our point-by-point response to each comment of the reviewers. Furthermore, I would like to take this opportunity to thank you for handling the review of our manuscript.

Our responses to the review comments are in blue.

Sincerely yours,
Ying Li, Ph.D.

College of Computer Science and Technology
Jilin University
Qianjin Street 2699, Changchun, Jilin 130012, P.R.China
Phone:  86-13504319660 (Mobile)

Response to Editor Comments

----------------------------------------------------------------------------------------------------------------------

**************
Editor Comments

Associate Editor
Comments to the Author:
This manuscript was reviewed by two experts.

Both of them have concerns on comparison with other methods, ways of computational experiments, and statistical tests.

Furthermore, one reviewer recommends that the type of the paper should be changed to regular one. And, I agree with this opinion.
(For page length/paper type issue, please do not ask me instead ask to the editorial staff or the editor in chief.)

Based on these points, I recommend the authors to revise the manuscript with taking all reviewers' comments into account.
*******************
Response: Thanks for providing us the opportunity to revise the manuscript. The revised version considering all remarks of the reviewers has just been submitted. We have substantially revised the previous manuscript and made great efforts in responding to the review comments. In addition, we have changed the type of the paper in the online system to a regular research one as you suggested. The response to each comment of the reviewers in the detail are provided as follows.

Response to Comments of Reviewer 1
---------------------------------------------------------------------------

Reviewer: 1

Recommendation: Author Should Prepare A Minor Revision

Comments:
A multilayer recursive feature elimination technique based on embedded genetic algorithm for cancer classification has been presented. The authors have proposed a hybrid technique comprising both filter and wrapper methods for gene subset selection. The work is interesting and the manuscript is well organized.

1. In the introduction section author has mentioned the phrase "lack an explicit decline of the feature number". The particular phrase is not clear. Please elaborate and explain clearly the lacuna of swarm intelligence based gene selection approach.

Response: Thank you for this comment. In the revised manuscript, we have added more explanations on the phrase "lack an explicit decline of the feature number" to account for the lacuna of swarm intelligence based gene selection approaches. 
•	The sentence containing the mentioned particular phrase in the "Introduction" section is:
"Nevertheless, all these feature selection methods based on swarm intelligence algorithms use the binary encoding method and lack an explicit reduction in the feature number."
•	We have added more explanations in the revised manuscript after the above sentence:
"The feature number only changes in the randomized evolution operation like mutation and crossover. Thus, these methods lack the precise control over the gene features in the individuals and can not explicitly remove genes to decrease the feature number."

2. In algorithm 1, it has been mentioned to sort the optimal gene combination in GC and to preserve the top ranked genes. On what basis the top ranked gene would be sorted?  For sorting what procedure is used?

Response: Thanks for your comment. For the first question, we sort the gene combinations based on two metrics: fitness and gene number. The individual with higher fitness is superior. For two individuals with the same fitness values, the one with a smaller gene number is superior. The fitness of an individual is defined respectively according to different datasets. On imbalanced datasets, the fitness is defined as α*Acc+(1-α)*Avc. On balanced datasets, the fitness is simply defined as accuracy Acc (more explanations of fitness definition are available in the section 2.3.2 “Stage 2: Precise wrapper search” in the revised manuscript).
For the sorting procedure, we use the TimSort [1] method, which is the default sorting procedure of Python: TimSort is a hybrid stable sorting algorithm integrated merge sorting and insertion sorting, which has O(n log n) time complexity for the worst case and O(n) for the best case scenarios. It performs well on the complex real-world data and can speed up the gene combination sorting process in our experiments.

3. In the search space reduction stage, it has been mentioned that top 1000 genes have been selected by a threshold of 0.05 in t-test technique and thereafter MIC has been applied on the 1000 genes to re-rank them. Is there any particular reason of selection 0.05 value as threshold? Is there any mathematical reason for selecting particularly this value for threshold in t-test? Or it has been selected experimentally and any other value can also be chosen? Clarify in detail.

Response: Thank you for this comment. The selection 0.05 value as threshold is on the basis of mathematical and statistically theories. It is worth noting that the standard level of significance used to justify a claim of a statistically significant effect is 0.05 and there are many theories to account for the use of 0.05 in denotation statistical significance, which can trace back to the influence of Statistical Methods for Research Workers by R.A. Fisher [2]. For better or worse, the term statistically significance has become synonymous with p≤ 0.05. Meanwhile, it is convenient to take the value of this threshold as a limit in judging whether a deviation ought to be considered significant or not. Overall, in the majority of analyses, an p of 0.05 is used as the cutoff for significance, guaranteeing the feasibility of most researches and studies.
   In addition, the reason why our proposed method MGRFE selected 0.05 as threshold in t-test technique is based on both theoretical and experimental foundations.
a)	The p=0.5 threshold is widely adopted in t-test among researchers owing to the characteristics of the selection 0.05 value mentioned above, forming a common standard in research result explain and comparison processes.
b)	The threshold of p=0.05 in t-test is experimentally selected based on the data distribution condition of 17 binary classification datasets and then validated by the satisfying experiment results on these datasets. Table 1 illustrates the number of significant genes with p-value less than 0.05 according to the t-test. From Table 1 we can note that, p=0.05 is a relatively accommodative condition on 17 binary-class datasets, which can not only identify the most differentially expressed genes, but also avoid the inappropriately exclusion of too many genes. This Table 1 becomes the Table 5 in Supplementary Material.

Table 1. Number of statistically significant features with t-test-based p-value less than 0.05 on 17 binary classification datasets.
Dataset	DLBCL	Pros	Colon	Leuk	Mye	ALL1	ALL2	ALL3	ALL4
Significant features	2632	5061	594	2449	1720	4387	644	571	1279
Total features	7129	12625	2000	7129	12625	12625	12625	12625	12625

Dataset	CNS	Lym	Adeno	Gas	Gas1	Gas2	T1D	Stroke	
Significant features	334	804	1799	8260	16454	15601	10159	5569	
Total features	7129	4026	7457	22645	22283	22283	54675	54675

4. What is the rationale for using particularly t-test first and then MIC? Can any combination of other two filter methods be used in search space reduction task? Clarify in detail. Use any other combination of two filter methods and compare it to the proposed combination of t-test and MIC-based search space reduction.

Response: Thanks for your comment.
(1)	T-test and MIC are used in gene filtering process due to their efficiency and convenience. The t-test has been widely used to identify differentially expressed genes in gene expression profiles [3-5]. But t-test has its limitation on dealing with multi-class dataset. The multi-variate t-test can't be performed directly. The recently proposed MIC shows excellent performance on detecting a wide range of associations in large datasets including microarray [6, 7], which can be used for multi-class dataset directly. Thus, we combined t-test and MIC to reduce the search space.
(2)	Using t-test first primarily because that the t-test can quickly reduce the gene number by the commonly used p=0.05 significance threshold. Compared to t-test, the calculation of MIC is more time-consuming. Therefore, it is reasonable to perform t-test first to quickly reduce the gene number, then use MIC to further select the significant gene on a relatively small search space. 
(3)	The other combination of two filter methods Anova then Fold change (FC) is used in the search space reduction task. We use the combination of first Anova then Fold change (FC) to compare with the combination of first t-test then MIC as shown in Table 2. The experiment was carried on 3 balanced datasets (Adeno, Gas1 and Pros) and 3 imbalanced datasets (DLBCL, Leuk, and CNS) using 5-fold cross validation. For Anova, the p-value threshold was also set as 0.05. From Table 2, for the combination of Anova+FC, the sizes of the finally selected genes are 2, 4, and 8 on datasets Adeno, Leuk and CNS, respectively. But for t-test+MIC combination used in our proposed MGRFE, the selected 1, 2 and 7 genes can achieve the same performance on the 3 datasets. On the rest of 3 datasets, the two combinations have similar performance. Thus, the filter method combination of Anova+FC is little inferior to the original combination of t-test+MIC in finding the minimal discriminative gene subset. Due to the length restriction of the main manuscript, we added the filter method combination results to the “S7” section in Supplementary Material. The Table 2 becomes part of the Table 6 in the supplementary section “S7”.

Table 2. Comparison of the combination of Anova and Fold change (FC) with the combination of the t-test and MIC on 3 balanced datasets (Adeno, Gas1 and Pros) and 3 imbalanced datasets (DLBCL, Leuk, and CNS) using 5-fold cross validation.
Filter Methods	Dataset	Genes	Sn	Sp	Acc	Avc	MCC	AUC
Anova+FC	Adeno	2	1	1	1	1	1	1
	Gas1	3	0.987	0.973	0.98	0.98	0.96	0.979
	Pros	4	0.98	0.982	0.981	0.981	0.963	0.982
	DLBCL	3	1	1	1	1	1	1
	Leuk	4	1	1	1	1	1	1
	CNS	8	1	1	1	1	1	1
t-test+MIC	Adeno	1	1	1	1	1	1	1
	Gas1	3	0.986	0.973	0.98	0.98	0.961	0.99
	Pros	4	0.98	0.982	0.981	0.981	0.963	0.98
	DLBCL	3	1	1	1	1	1	1
	Leuk	2	1	1	1	1	1	1
	CNS	7	1	1	1	1	1	1

5. The researchers have used t-test and then MIC. The gene selected after MIC is used in the proposed MGRFE algorithm. In table 5, why the t-test-based gene ranking has been compared? MIC based ranking should also be compared.

Response: Thanks for your suggestion. In the Table 1 in our previous manuscript, we listed the t-test ranking results for all selected genes on 17 binary datasets (is table 5 a mistake of Table 1 ?). In the revised manuscript, we have added the MIC-based gene rankings to the Table 1 and the results explanation are also added in the subsection 3.1 “Results on Dataset One”. For convenience to review, the t-test and MIC-based gene ranking results are listed in Table 3. 
From the Table 2, it could be noted that: a) The relative positions of selected genes in the two ranking methods are consistent on the most datasets. For example, on the DLBCL dataset, the selected genes are ranked [13, 39, 54] in the t-test sorting. Meanwhile, the MIC-based gene rankings for same genes are [8, 24, 52], keeping the same ascending order as in t-test. b) The top-ranked genes in the t-test are also top-ranked in the MIC ranking. For example, the selected gene on dataset ALL1 is the top one (with ranking 1) in both t-test and MIC sorting process. c) The selected discriminatory genes are usually top-ranked by the t-test and MIC methods. For 5 of 17 datasets, the top one gene according to the t-test appeared in the final selected gene subsets. The t-test and MIC could find the informative genes, which are important for the later feature wrapper search process. Therefore, the employed filter techniques are qualified for the search space reduction stage.

Table 3. The sizes of selected gene subsets for t-test and MIC-based gene ranking results on the 17 binary classification datasets.

Datasets	Genes	t-test/MIC-based gene rankings
DLBCL	3/7129	[13/8, 39/24, 54/52]
Pros	4/12625	[1/1, 15/47, 74/49, 694/618]
Colon	6/2000	[15/6, 58/21, 176/297, 225/80, 240/555, 495/482]
Leuk	2/7129	[4/3, 7/5]
Mye	7/12625	[3/3, 15/103, 83/142, 143/13, 378/217, 404/644, 569/707]
ALL1	1/12625	[1/1]
ALL2	8/12625	[1/80, 52/395, 78/3040, 80/1297, 522/2448, 687/2038, 737/920, 760/1449]
ALL3	8/12625	[4/500, 52/3437, 75/3010, 142/393, 488/443, 510/795, 715/1551, 770/1321]
ALL4	6/12625	[1/2, 6/45, 39/356, 282/226, 535/497, 754/1377]
CNS	7/7129	[9/907, 53/542, 130/620, 131/519, 272/57, 273/454, 520/49]
Lym	3/4026	[4/7, 5/4, 669/135]
Adeno	1/7457	[468/27]
Gas	3/22645	[22/1, 77/32, 306/36]
Gas1	3/22283	[132/74, 248/167, 717/500]
Gas2	2/22283	[38/6, 89/62]
T1D	7/54675	[14/2229, 25/1579, 113/1287, 559/1282, 578/353, 680/426, 978/1728]
Stroke	4/54675	[1/3, 23/115, 129/543, 276/539]
In the “Genes” column, we list the number of selected gene among the total genes in each dataset. In the “t-test/MIC-based gene rankings” column, we list the ranking positions for each gene in the t-test and MIC processes.

6. Elaborate the significance of '0' ranked gene in t-test.

Response: Thank you for this comment. For the index of the first element in a general array in common programming language is '0', so in our previous manuscript, the value '0' is just assigned to the first gene in the t-test gene ranking result.
In the revised manuscript, we have changed the initial index to '1' in order to be understood clearly.

7. The comparative results of SRBCT, ALL-AML and ALL have been shown in table 7, 8 and 9. However, the tables are very similar to work in kar et al. [28]. The similar type of comparison should be given for all the dataset used i.e. 19 dataset in the present work.

Response: Thanks for your comment. The chief reason why we have based on the tables of Kar et al. to made comparison of results is that their tables have been well organized, readable and understandable. In fact, we have added several new items on the tables in the work by Kar et al., so that it is easy to better evaluate the effectiveness on those three datasets located in previous work and our proposed method.
   For similar comparison of other datasets, the results are consistent. Therefore, in the revised manuscript, we have added two typical selected gene comparison on DLBCL and Prostate.

Dataset	Criteria	EPSO[8] IBPSO[9] TS-BPSO[10] BPSO-CGA[11] Random forest	mABC[12] MGRFE
DLBCL	Acc	100	100	100	-	94.6	100	100
		#Genes	3	1042	2671	-	21	3	3
Pros	Acc	99.02	92.16	95.45	93.7	-	100	98.1
		#Genes	5	1294	5320	795	-	5	4

8. The proposed work has also been compared with Kar et al. [28] in computational performance. Kar et al. have applied a swarm intelligence-based method to the space of all genes. They have not reduced the search space prior to the optimization task. In contrast, the proposed method have applied MGRFE technique on the reduced search space. The reduce search space have been constructed by t-test and then by MIC technique. In my opinion the search space reduction is fixed. It is done once before the application of MGRFE. In that regard, the comparison of computational time would not significant because it has been computed in the reduced search space. The genes outside the reduced search space could carry valuable information towards classification accuracy.

Response: Thank you for this comment.
1)	Indeed, the lack of filter preprocessing could increase the running time of Kar et al.’s method. But it should be noticed that the recursive feature elimination (RFE) manner dramatically can reduce the running time of MGRFE. Both the method by Kar et al. and our proposed algorithm MGRFE employed the time-consuming swarm intelligence methods (i.e. PSO and GA).But we combined the GA with RFE,which notably speeded up the convergence process. It should also be pointed out that the computation time of MGRFE includes the whole filter screen and MGRFE wrapper search process, rather than on a fixed reduced feature search space. For example the multi-class SRBCT dataset, with 2308 genes in total, Kar et al.’s PSO+KNN method 2.7956 hours to select the final gene subset. However, MGRFE merely demanded 10.828 minutes totally (including whole filter and MGRFE wrapper process) to select the final gene subset, about 6.46% of the time spent by Kar et al.’s method. Besides, the 5 genes selected by MGRFE obtained higher classification accuracy than the 6 genes finally selected by Kar et al’s method on dataset SRBCT.
2)	For the filter process has already supplied sufficient discriminatory expressed genes for finding the minimal discriminatory gene subset, MGRFE just ignore the rest genes outside the reduced feature space. Our filter process using the combination of t-test and MIC selects significant genes related to the phenotypes which are very valuable for the subsequent feature selection process. There are several thousand to tens of thousands of genes in a microarray, among which the most are irrelevant features. If take all gene features into consideration, the irrelevant and redundant features will disturb the wrapper algorithm to select the minimal informative gene subset with higher time cost.

9. In the Conclusion section, the authors will need to clearly address the research contributions in theory. The research contributions in theory must be fully stated in at least one paragraph.

Response: Thanks for your comment. To better illustrate the contributions of our algorithm, we have added more detailed statements in the section of discussion and conclusion in the revised manuscript. The primary research contribution in theory is to provide an effective feature selection method with low time cost which combines the embedded genetic algorithm with recursive feature elimination process. To the best of our knowledge, none previous studies have been designed as an evolutionary algorithm using variable length integer encoding approach in a recursive process to deal with the problem of minimal discriminatory feature selection for high-dimension datasets. Meanwhile, through the experimental comparisons with the popular feature selection on diverse datasets, our proposed MGRFE could outperform the most other state-of-the-art algorithms for gene selection on microarray data.Therefore,MGRFE is a novel effective and efficient feature selection alyorithm to offer the better options for users.

10. In the Conclusion section, the authors need to fully discuss insightful and practical implications.

Response: Thank you for this comment. We have supplemented sufficient practical significance and applications to make the our proposed MGRFE more convincing and valuable. The proposed MGRFE can be regarded as a promising method to the potential cancer diagnosis by selecting the most related and minimal discriminatory genes, and it can aslo be widely applied to the analyses of high-throughput genomic, proteomic and metabolomics data. Moreover, the biological association between selected gene subsets with the related cancer phenotypes are validated by the literature mining on PubMed, which could provide novel cancer biomarkers candidates for clinic researches.

Response to Comments of Reviewer 2
---------------------------------------------------------------------------

Reviewer: 2

Recommendation: Author Should Prepare A Major Revision For A Second Review

Comments:
1.	First of all, the paper is described as "Survey/Tutorial," but it appears to describe a claimed original contribution by the authors, namely the MGRFE algorithm. The proposed new algorithm is compared against several existing algorithms. Therefore, if at all the paper is to be published, it should be as a regular research paper, and not as a survey/tutorial paper.

Response: Thank you for this comment. We have changed the type of the paper to a regular research one in the online system.

2.	The paper is a mixture of techniques that are by now standard in the world of computational biology.  Given a very large number of features, first use some pre-filtering to eliminate perhaps 90% to 95% of the features, and then use recursive feature elimination (RFE) on the remaining features.  I could not find any compelling evidence that the proposed approach is superior to the existing methods.

### 2018-12-01 MGRFE Response 2018-12-01/Latex Manuscript/MGRFE-tcbb.tex
Complete! Record my gratitude.

	Wrapper algorithms usually employ classification models and contain heuristic rules to select feature subsets guided by the classification performance on the feature subsets being used, which are usually superior to filter algorithms but more time-consuming. A variety of wrapper algorithms have emerged involving simulated annealing, randomized hill climbing \cite{W26}, regularized random forest (RRF) \cite{W11}, particle swarm optimization (PSO) \cite{W12,W10}, and genetic algorithm (GA) \cite{W13}. With the rapid development of heuristic rules and evolutionary strategies commonly present in wrapper techniques, various swarm intelligence algorithms have been applied to the optimization of feature selection. Kar \emph{et~al.} have proposed a particle swarm optimization method based on adaptive K-nearest neighborhood (KNN) to identify a minimum meaningful gene subset \cite{W15}. Moosa \emph{et~al.} have presented a modified artificial bee colony algorithm (ABC) to select a minimum number of genes with high predictive accuracy for cancer classification \cite{P32}. Oreski \emph{et~al.} have designed a hybrid GA with neural networks to identify an optimal feature subset with high classification accuracy and scalability for credit risk assessment \cite{W17}. Jung and Zscheischler have described a guided hybrid GA to minimize the number of cost function evaluations \cite{W18}. Nevertheless, all these feature selection methods based on swarm intelligence algorithms use the binary encoding method and lack an explicit reduction in the feature number. Meanwhile, it has been verified that only a minimal number of informative genes is enough for effective diagnosis of different phenotypes in microarray gene datasets \cite{P1,W16,P32,P33}. The feature selection using binary encoding has three main shortcomings in finding an optimal gene combination: (1) The fixed chromosome length for the encoding length must be equal to the gene range to represent all the genes. This arrangement can result in impossibility of the explicit reduction in the gene number and unnecessary space occupation when there are only several 1s among lots of 0s. (2) There are different numbers of actual existing genes in different individuals. Because there are different numbers of 1s, the actual number of genes varies among individuals and cannot be controlled precisely. (3) The convergence speed is usually low and the time cost is high to generate the minimal informative gene combination. 
	The sizes of the optimal gene combinations in most of datasets are below 10. The evolution-based feature selection algorithms using binary encoding lack of an explicit feature reduction mechanism, which results in low probability and high time cost to generate the optimal minimal gene combination among the several thousand to tens of thousands of genes in each dataset.
	Recursive feature elimination (RFE) is a popular strategy that yields an explicit recursive feature reduction by removing features with the least weights \cite{W7,P1,P2,P27}.

@@

	\begin{figure*}[!htbp]
		\centering
		\includegraphics[width=0.68\textwidth]{pictures/MGRFE-whole.pdf}
		\caption{A flowchart of the whole MGRFE procedure in this study.}
		\label{Fig:MGRFE_whole}
	\end{figure*}

@@

	{\footnotesize
		\begin{equation}\label{eq:1}
		\begin{split}
		Sn = & \frac{TP}{TP+FN} ,\quad Sp = \frac{TN}{TN+FP}, \\
		Acc = & \frac{TP+TN}{P+N} ,\quad Avc = \frac{Sn+Sp}{2}, \\
		MCC = & \frac{TP \cdot TN-FP \cdot FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}\\
		\end{split}
		\end{equation}
	}

@@

	\begin{figure*}[t]
		\centering
		\includegraphics[width=0.9\textwidth]{pictures/MGRFE-flowchart.pdf}
		\caption{The flowchart of MGRFE, which is divided into 3 stages: search space reduction, precise wrapper search, and multiple k-fold CV. Stage 2 is the core of MGRFE, which includes two key processes: GaRFE and embedded modified GA.}
		\label{Fig:MGRFE_flowchart}
	\end{figure*}

@@

	At Stage 1, two filter methods, the \textit{t}-test and MIC, are used to decrease the gene range and offer a candidate gene set for later precise wrapper search stage. First, we perform the \textit{t}-test on all genes and subject them to ascending sorting according to their $p$ values, then the top-ranked significant features with $p$ values less than 0.05 are preserved. Next, the upper limit of the features kept after the \textit{t}-test is set to 1000, that is, when there are more than 1000 features having $p$ values less than 0.05, only the top 1000 with lower $p$ values would be kept. If the preserved features after the \textit{t}-test screening are fewer than 500, they are all kept directly and definitively to form the candidate gene set without MIC screening; otherwise, the MIC-based selection will be followed. Second, we carry out MIC calculation on the preserved genes and resort them according to their MIC values, then the candidate gene set is generated from the top 500 genes with higher MIC values. For the two multiclass datasets, a candidate gene set is generated based only on the descending order of MIC values of all genes for which the multivariate \textit{t}-test cannot be performed directly.

@@

	The results of MGRFE on Dataset One including 17 binary datasets are given in Table \ref{Tab:RST all index}, where six measurements calculated by 5-fold CV and the \textit{t}-test and MIC-based gene rankings are listed. For 17 datasets, \(Acc\) values are all above 0.9 within 10 genes. Moreover, for 8 of 17 datasets (DLBCL, Leuk, ALL1, Lym, Adeno, Gas, Gas2, and Stroke), \(Acc\) reached 1.0 with gene number less than 5. MGRFE also show the strong robustness in dealing with imbalanced datasets like DLBCL, Colon, Leuk, ALL1, ALL4, and CNS, for which \(Sn\), \(Sp\), \(Avc\), \(MCC\), and \(AUC\) are all above 0.95 without being influenced by the data imbalance. 
	According to the \textit{t}-test and MIC-based gene ranking, the best gene feature subset is not always the highest-ranked features in the filter method, thus the filter algorithm alone cannot generate the optimal feature combination.
	It could be noted that the relative positions of selected genes in the two ranking methods are consistent on most datasets. The top-ranked genes in the \textit{t}-test are also top-ranked in the MIC sorting (e.g. the selected gene on ALL1 is the top one in both \textit{t}-test and MIC ranking). For 5 of 17 datasets, the top one gene  according to the \textit{t}-test appeared in the final selected gene subsets. Generally, the selected informative genes are top-ranked by the \textit{t}-test and MIC methods. Therefore, the filter techniques are qualified for the search space reduction stage. Moreover, MGRFE achieves stable classification performance in 10 repetitions of 10-fold CV as depicted in Fig. \ref{Fig:CV1}.

>>>>

	Wrapper algorithms usually employ classification models and contain heuristic rules to select feature subsets guided by the classification performance on the feature subsets being used, which are usually superior to filter algorithms but more time-consuming. A variety of wrapper algorithms have emerged involving simulated annealing, randomized hill climbing \cite{W26}, regularized random forest (RRF) \cite{W11}, particle swarm optimization (PSO) \cite{W12,W10}, and genetic algorithm (GA) \cite{W13}. With the rapid development of heuristic rules and evolutionary strategies commonly present in wrapper techniques, various swarm intelligence algorithms have been applied to the optimization of feature selection. Kar \emph{et~al.} have proposed a particle swarm optimization method based on adaptive K-nearest neighborhood (KNN) to identify a minimum meaningful gene subset \cite{W15}. Moosa \emph{et~al.} have presented a modified artificial bee colony algorithm (ABC) to select a minimum number of genes with high predictive accuracy for cancer classification \cite{P32}. Oreski \emph{et~al.} have designed a hybrid GA with neural networks to identify an optimal feature subset with high classification accuracy and scalability for credit risk assessment \cite{W17}. Jung and Zscheischler have described a guided hybrid GA to minimize the number of cost function evaluations \cite{W18}. Nevertheless, all these feature selection methods based on swarm intelligence algorithms use the binary encoding method and lack an explicit reduction in the feature number. The feature number only changes in the randomized evolution operation like mutation and crossover. Thus, these methods lack the precise control over the gene features in the individuals and can not explicitly remove genes to decrease the feature number. Meanwhile, it has been verified that only a minimal number of informative genes is enough for effective diagnosis of different phenotypes in microarray gene datasets \cite{P1,W16,P32,P33}. The feature selection using binary encoding has three main shortcomings in finding an optimal gene combination: (1) The fixed chromosome length for the encoding length must be equal to the gene range to represent all the genes. This arrangement can result in impossibility of the explicit reduction in the gene number and unnecessary space occupation when there are only several 1s among lots of 0s. (2) There are different numbers of actual existing genes in different individuals. Because there are different numbers of 1s, the actual number of genes varies among individuals and cannot be controlled precisely. (3) The convergence speed is usually low and the time cost is high to generate the minimal informative gene combination. 
	The sizes of the optimal gene combinations in most of datasets are below 10. The evolution-based feature selection algorithms using binary encoding lack of an explicit feature reduction mechanism, which results in low probability and high time cost to generate the optimal minimal gene combination among the several thousand to tens of thousands of genes in each dataset.
	Recursive feature elimination (RFE) is a popular strategy that yields an explicit recursive feature reduction by removing features with the least weights \cite{W7,P1,P2,P27}.

@@

	\begin{figure*}[!htbp]
		\centering
		\includegraphics[width=0.6\textwidth]{pictures/MGRFE-whole.pdf}
		\caption{A flowchart of the whole MGRFE procedure in this study.}
		\label{Fig:MGRFE_whole}
	\end{figure*}

@@

	{\scriptsize
		\begin{equation}\label{eq:1}
		\begin{split}
		Sn = & \frac{TP}{TP+FN} ,\quad Sp = \frac{TN}{TN+FP}, \\
		Acc = & \frac{TP+TN}{P+N} ,\quad Avc = \frac{Sn+Sp}{2}, \\
		MCC = & \frac{TP \cdot TN-FP \cdot FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}\\
		\end{split}
		\end{equation}
	}

@@

	\begin{figure*}[t]
		\centering
		\includegraphics[width=0.87\textwidth]{pictures/MGRFE-flowchart.pdf}
		\caption{The flowchart of MGRFE, which is divided into 3 stages: search space reduction, precise wrapper search, and multiple k-fold CV. Stage 2 is the core of MGRFE, which includes two key processes: GaRFE and embedded modified GA.}
		\label{Fig:MGRFE_flowchart}
	\end{figure*}

@@

	At Stage 1, two filter methods, the \textit{t}-test and MIC, are used to decrease the gene range and offer a candidate gene set for later precise wrapper search stage. First, we perform the \textit{t}-test on all genes and subject them to ascending sorting according to their $p$ values, then the top-ranked statistically significant features with $p$ values less than 0.05 are preserved. Next, the upper limit of the features kept after the \textit{t}-test is set to 1000, that is, when there are more than 1000 features having $p$ values less than 0.05, only the top 1000 with lower $p$ values would be kept. If the preserved features after the \textit{t}-test screening are fewer than 500, they are all kept directly and definitively to form the candidate gene set without MIC screening; otherwise, the MIC-based selection will be followed. Second, we carry out MIC calculation on the preserved genes and resort them according to their MIC values, then the candidate gene set is generated from the top 500 genes with higher MIC values. For the two multiclass datasets, a candidate gene set is generated based only on the descending order of MIC values of all genes for which the multivariate \textit{t}-test cannot be performed directly. 
	In the Table 5 in Supplementary Material, the number of statistically significant genes with \emph{t}-test-based \emph{p}-values less than 0.05 on each binary-class dataset are listed.
	In the "S8" section of Supplementary Material, we also give a simple comparison of \textit{t}-test+MIC with other filter combinations.

@@

	The results of MGRFE on Dataset One including 17 binary datasets are given in Table \ref{Tab:RST all index}, where six measurements calculated by 5-fold CV and the \textit{t}-test and MIC-based gene rankings are listed. For 17 datasets, \(Acc\) values are all above 0.9 within 10 genes. Moreover, for 8 of 17 datasets (DLBCL, Leuk, ALL1, Lym, Adeno, Gas, Gas2, and Stroke), \(Acc\) reached 1.0 with gene number less than 5. MGRFE also show the strong robustness in dealing with imbalanced datasets like DLBCL, Colon, Leuk, ALL1, ALL4, and CNS, for which \(Sn\), \(Sp\), \(Avc\), \(MCC\), and \(AUC\) are all above 0.95 without being influenced by the data imbalance. 
	According to the \textit{t}-test and MIC-based gene ranking, the best gene feature subset is not always the highest-ranked features in the filter method, thus the filter algorithm alone cannot generate the optimal feature combination.
	It could be noted that the relative positions of selected genes in the two ranking methods are consistent on most datasets. The top-ranked genes in the \textit{t}-test are also top-ranked in the MIC sorting (e.g. the selected gene on ALL1 is the top one in both \textit{t}-test and MIC ranking). For 5 of 17 datasets, the top one gene  according to the \textit{t}-test appeared in the final selected gene subsets. Generally, the selected informative genes are top-ranked by the \textit{t}-test and MIC methods. Therefore, the filter techniques are qualified for the search space reduction stage. Moreover, MGRFE achieves stable classification performance in 10 repetitions of 10-fold CV as depicted in Fig. \ref{Fig:CV1}. Besides, we also validated the selected gene features of Leuk, Gas1 and Gas2 on independent datasets in the "S8" section of Supplementary Material.

	\begin{table*}[!htbp] \footnotesize
		\centering
		\caption{Results of MGRFE on 17 datasets in Dataset One}
		\label{Tab:RST all index}
		\resizebox{0.95\textwidth}{!}{
			\begin{tabular}{l|lllllllll}
				\toprule
				Datasets & Pos/Neg & Genes/Total & $Sn$ & $Sp$ & $Acc$ & $Avc$ & $MCC$ & $AUC$ & \textit{t}-test/MIC-based gene rankings\\
				\midrule
				DLBCL & 58/19 & \textbf{3}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [13/8, 39/24, 54/52]\\
				Pros & 52/50 & \textbf{4}/12625 & 0.980 & 0.982 & \textbf{0.981} & 0.981 & 0.963 & 0.98 & [1/1, 15/47, 74/49, 694/618]\\
				Colon & 40/22 & \textbf{6}/2000 & 1.0 & 0.960 & \textbf{0.985} & 0.980 & 0.969 & 0.97 & [15/6, 58/21, 176/297, 225/80, 240/555, 495/482]\\
				Leuk & 47/25 & \textbf{2}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [4/3, 7/5]\\
				Mye & 137/36 & \textbf{7}/12625 & 0.963 & 0.839 & \textbf{0.937} & 0.901 & 0.816 & 0.95 & [3/3, 15/103, 83/142, 143/13, 378/217, 404/644, 569/707]\\
				ALL1 & 95/33 & \textbf{1}/12625 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [1/1]\\
				ALL2 & 65/35 & \textbf{8}/12625 & 0.914 & 0.908 & \textbf{0.910} & 0.911 & 0.829 & 0.94 & [1/80, 52/395, 78/3040, 80/1297, 522/2448, 687/2038, 737/920, 760/1449]\\
				ALL3 & 24/101 & \textbf{8}/12625 & 0.830 & 0.950 & \textbf{0.927} & 0.890 & 0.785 & 0.93 & [4/500, 52/3437, 75/3010, 142/393, 488/443, 510/795, 715/1551, 770/1321]\\
				ALL4 & 26/67 & \textbf{6}/12625 & 1.0 & 0.986 & \textbf{0.990} & 0.993 & 0.978 & 0.99 & [1/2, 6/45, 39/356, 282/226, 535/497, 754/1377]\\
				CNS & 39/21 & \textbf{7}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 0.98 & [9/907, 53/542, 130/620, 131/519, 272/57, 273/454, 520/49]\\
				Lym & 22/23 & \textbf{3}/4026 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [4/7, 5/4, 669/135]\\
				Adeno & 18/18 & \textbf{1}/7457 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [468/27]\\
				Gas & 29/36 & \textbf{3}/22645 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [22/1, 77/32, 306/36]\\
				Gas1 & 72/72 & \textbf{3}/22283 & 0.986 & 0.973 & \textbf{0.980} & 0.980 & 0.961 & 0.99 & [132/74, 248/167, 717/500]\\
				Gas2 & 62/62 & \textbf{2}/22283 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [38/6, 89/62]\\
				T1D & 57/44 & \textbf{7}/54675 & 0.911 & 0.912 & \textbf{0.911} & 0.912 & 0.826 & 0.94 & [14/2229, 25/1579, 113/1287, 559/1282, 578/353, 680/426, 978/1728]\\
				Stroke & 20/20 & \textbf{4}/54675 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [1/3, 23/115, 129/543, 276/539]\\
				\bottomrule
		\end{tabular}}
	\end{table*}
	\begin{figure*}[htbp]
		\centering
		\includegraphics[width=0.72\textwidth]{pictures/CV1.pdf}
		\caption{The distribution of $Acc$ values in 10-time 10-fold CV for the selected gene combinations of 17 datasets in Dataset One.}
		\label{Fig:CV1}
	\end{figure*}
